---
title: Hello reprex
editor_options: 
  chunk_output_type: console
author: ''
date: '2020-10-29'
slug: hello_reprex.en-us
categories:
  - howto
tags:
  - data science
  - SAS
  - Python
  - R
keywords:
  - methods
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

*Disclaimer* : I am a huge fan of the R language thus this post is biased toward the few things I know about Python and SAS linguas.


## Code practice

One of the first advice I would give my past stat student self is to get better at coding ASAP. Indeed knowing better a language shapes also your thinking so you can organize yourself better with the task at hand and not feel too overwhelmed (like during your internship).

There are many ways to practice your coding skill such as doing data analysis project (e.g with [TidyTuesday](https://github.com/rfordatascience/tidytuesday)), modeling project ([kaggle](https://www.kaggle.com/) or [challengedata](https://challengedata.ens.fr/challenges/year/2020)), contributing to open source (by fixing github issues) or trying to help people with their questions on forums.

We will focus today on what I think worked for me : trying to help on forums. There are several websites you can choose from including :

  * [StackOverflow](https://stackoverflow.com/)
  * Language specific community site
    * [SAS communities](https://communities.sas.com/)
    * [RStudio community](https://community.rstudio.com/)
  * Open source language forums and mailing list.
    * [python.org](https://www.python.org/)
    * [r-project.org](https://www.r-project.org/)
  * Google groups
  * ...


## You said REPRoducible EXemple (REPREX) ?

I first came to know of **reproducible example** while practicing coding with `Stack Overflow`. There are two simple and difficult subjects when interacting on site and forums in **asking a good question** and **giving a good answer**. Indeed asking or trying to help :

  * compels you to sum up whatever problems you are trying to solve into (ideally) smaller code problems. This is important since you will get clear output : a well formulated code problem will either work or not.
  
  * to that end you have to share sufficient sample data (thus the **example**) and runnable code (hence the **reproducible**) to illustrate what you want to achieve.
  
In that perspective `StackOverflow` is a good place to practice and learn with enough gamification with points and bronze/silver/gold badges to keep you motivated for a time and feedback (sometime nice, sometime harsh) on the quality of your contribution.
  

**To sum up** : a **REPRodicible EXample** is a bundle of information (data, runnable codes, environment, OS and package version) allowing one to share his results and error identically.

Such object is quite useful to :

  * ensure that someone gets to the point where you are and understand what you want to achieve [question side] 
  * reproduce error messages and bugs [general side]
  * share a solution that should work for everyone [answer side]


SO Resources :

- [how-to-ask](https://stackoverflow.com/help/how-to-ask)
- [minimal-reproducible-example](https://stackoverflow.com/help/minimal-reproducible-example)

---

Today I was working with Python and SAS. Having used mostly `R` before, the first thing I compared was how to get help from the documentation (with examples, not always runnable) and forums (with reprex).

For example suppose you have to clean some data by replacing the ▒ character with something. One universal format used in data science software is the `data frame` (Python, R) or the `table` (SAS) data format. Those objects allows one to wrangle data and organize it in a meaningful way for analysis. 

```{r create data, echo = FALSE}
# table example here
set.seed(456)
dfr <- tibble::tibble(
  var1 = sample(stringr::fruit, 6),
  var2 = sample(stringr::fruit, 6),
  var3 = sample(1:6, 6, replace = TRUE)
)
dfr$var1 <- stringr::str_replace_all(string = dfr$var1, pattern =  "a", replacement = "▒")
dfr$var2 <- stringr::str_replace_all(string = dfr$var2, pattern =  "e", replacement = "▒")

knitr::kable(dfr, format = "simple")
```

---

### SAS

The first step to dive into this problem is to reproduce the data. In SAS there are several ways to share sample data such as :

* using a built-in sample data available (in the sashelp)

```{}
data work.test ;
  set sashelp.cars (obs = 10) ; 
run ;
```

* use a CSV file (or any popular other format) - this method is not recommended though since it puts extra work on the side of the one that wants to help you or reproduce your research

```{}
/* use the work library for tests */
proc import file = "path_to_file" out = work.table_name ; run;
```

* with a data step and `datalines` command (I have cheated a little by adding the `|` separator to simplify things)

```{}
data test;
	infile datalines dsd dlm='|' truncover ;
	input var1 :$20. var2 :$20. var3 ;
	datalines ;
gr¦pefruit|ugli fruit|4
huckleberry|kumquat|6
duri¦n|blackcurrant|1
d¦te|cantaloup¦|1
fig|canary m¦lon|6
s¦l¦l berry|rambutan|5
;;;;
run;

/* that gives */
/*
> 1	gr¦pefruit	ugli fruit	4
> 2	huckleberry	kumquat	6	
> 3	duri¦n	blackcurrant	1	
> 4	d¦te	cantaloup¦	1	
> 5	fig	canary m¦lon	6	
> 6	s¦l¦l berry	rambutan	5
*/
```


They are many ways to fix this data with for the examples the use of the `prxchange` and the `tranwrd` functions :

```{}
data test_clean ;
	set test ;
	var1 = prxchange("s/¦/a/i", -1, var1) ;
	var2 = tranwrd(var2, "¦", "e") ;
run ;

/* result */
/*
> 1	grapefruit	ugli fruit	4	
> 2	huckleberry	kumquat	6	
> 3	durian	blackcurrant	1	
> 4	date	cantaloupe	1	
> 5	fig	canary melon	6	
> 6	salal berry	rambutan	5
*/
```

This concludes the resolution of the problem with a reprex in question (for sharing the data) and a reprex in solution that anyone having an installation of SAS 9.4 available should be able to run with copy/paste.


Note : aside from Stack Overflow, SAS users should find useful contents with the [SAS Communities](https://communities.sas.com/) website that is a language specific free help platform.

More ways to share SAS reprex with this [SAS Communities question](https://communities.sas.com/t5/SAS-Communities-Library/How-to-create-a-data-step-version-of-your-data-AKA-generate/ta-p/258712)


---


### Python

`Pandas` should be your go to when handling data with Python. Here to reproduce the data we can directly as a pandas.Data.Frame one can load the pandas package

```{r set up python version, include=FALSE}
# ref https://rstudio.github.io/reticulate/articles/versions.html
# reticulate::py_discover_config() // ne trouve pas la version voulue

# KO
# reticulate::use_python("C:\\Users\\cbonamy\\AppData\\Local\\Programs\\Python\\Python38-32")

# reticulate::use_python("C:\\Users\\cbonamy\\Anaconda3\\bin")

# en default de cause
# reticulate::py_install("pandas")
```


```{python pandas import}
import pandas as pd
```

then define the dataframe from its rows with 6 vectors :

```{python vector data frame}
dfr_vec = pd.DataFrame([ ["gr¦pefruit", "ugli fruit", 4],
				["huckleberry", "kumquat", 6],
				["duri¦n", "blackcurrant", 1],
				["d¦te", "cantaloup¦", 1],
				["fig", "canary m¦lon", 6],
				["s¦l¦l berry", "rambutan", 5] ],
			columns = ["var1", "var2", "var3"]
			)
dfr_vec
```

or using the dictionary from with :

```{python dictionary data frame}
dfr_dic = pd.DataFrame({"var1" : ["gr¦pefruit", "huckleberry","duri¦n","d¦te","fig","s¦l¦l berry"],
"var2" : ["ugli fruit","kumquat", "blackcurrant","cantaloup¦","canary m¦lon", "rambutan"],
"var3" : [4, 6, 1, 1, 6, 5]})
dfr_dic
```


* pd.read_clipboard : allows one to import the data as a  `pandas.DataFrame` from what is copied in the clipboard. Here for purpose of the example you can test it with copying the `print` (the `head` method won't work well) of the previous data frame :

```
          var1          var2  var3
0   gr¦pefruit    ugli fruit     4
1  huckleberry       kumquat     6
2       duri¦n  blackcurrant     1
3         d¦te    cantaloup¦     1
4          fig  canary m¦lon     6
5  s¦l¦l berry      rambutan     5
```

becomes :

```{}
# read data frame from clipboard
dfr_bis = pd.read_clipboard(sep='\s\s+')
dfr_bis

          var1          var2  var3
0   gr¦pefruit    ugli fruit     4
1  huckleberry       kumquat     6
2       duri¦n  blackcurrant     1
3         d¦te    cantaloup¦     1
4          fig  canary m¦lon     6
5  s¦l¦l berry      rambutan     5
```


Fixing the character issue with pandas :

```{python py fix string}
dfr_dic['var1'] = dfr_dic['var1'].str.replace('¦','a')
dfr_dic['var2'] = dfr_dic['var2'].str.replace('¦','e')
dfr_dic
```

More resources :
- [how-to-make-good-reproducible-pandas-examples](https://stackoverflow.com/questions/20109391/how-to-make-good-reproducible-pandas-examples)

- [pandas cheatcheet](https://pandas.pydata.org/Pandas_Cheat_Sheet.pdf)


---


### R

**Sharing data**

Most of the time the data already exists in memory (here with `dfr`). So you should sample it if needed 
with :

```{r r sample data}
dfr_head <- head(dfr)

# or randomly

dfr_sample <- dfr[ sample(x = 1:nrow(dfr), size = 6) , ]
```

Then you can either write yourself the data to share as below by rows or by columns or `dput()` the package `datapasta` to produce such output.

0. `dput` is a base R function to ouput "text representation of an R object"

```{r dput}
dput(dfr)
```


1. `datapaste::trible_paste` for printing inline an existing dataframe by rows

```{}
datapasta::tribble_paste(dfr)

tibble::tribble(
                   ~var1,          ~var2, ~var3,
            "gr¦pefruit",   "ugli fruit",    4L,
           "huckleberry",      "kumquat",    6L,
                "duri¦n", "blackcurrant",    1L,
                  "d¦te",   "cantaloup¦",    1L,
                   "fig", "canary m¦lon",    6L,
           "s¦l¦l berry",     "rambutan",    5L
           )
```

2. `datapasta::df_paste` for printing inline an existing dataframe by columns

```{}
datapasta::df_paste(dfr)

data.frame(
  stringsAsFactors = FALSE,
                 var1 = c("gr¦pefruit",
                          "huckleberry","duri¦n","d¦te","fig","s¦l¦l berry"),
                 var2 = c("ugli fruit","kumquat",
                          "blackcurrant","cantaloup¦","canary m¦lon",
                          "rambutan"),
              var3 = c(4L, 6L, 1L, 1L, 6L, 5L)
   )
```

3. From the clipboard you can copy the table below and paste is follows

```{r data to copy to clipboard, echo = FALSE}
# datapasta::tribble_paste(dfr)
knitr::kable( tibble::tribble(
                             ~var1,          ~var2, ~var3,
                      "gr¦pefruit",   "ugli fruit",    4L,
                     "huckleberry",      "kumquat",    6L,
                          "duri¦n", "blackcurrant",    1L,
                            "d¦te",   "cantaloup¦",    1L,
                             "fig", "canary m¦lon",    6L,
                     "s¦l¦l berry",     "rambutan",    5L
                     )
)
```

```{r datapaste tribble, eval=FALSE}
datapasta::tribble_paste()
```


**Answering the problem at hand and sharing your rersult in a reproducible way**

```{r r solutions, warning=FALSE, message=FALSE}
# Method 1 - base R
dfr_clean <- dfr
dfr_clean$var1 <- gsub(pattern = "¦", replacement = "a", x = dfr$var1)
dfr_clean[["var2"]] <- gsub(pattern = "¦", replacement = "e", x = dfr$var2)

# Method 2 - tidyverse
library(dplyr)
library(stringr)
dfr %>% mutate(var1 = str_replace_all(string = var1, pattern = "¦", replacement = "a"),
	       var2 = str_replace_all(string = var2, pattern = "¦", replacement = "e")
	       )
```


** Finally : Why sharing runnable sample code with R is easy **

Because some genius already made the hard work for you !

The package `reprex` was originaly created by Jenny Bryan. It allows one to ensure that you are effectively sharing a reprex as such :

  * if your code is not runnable in a new blank R session (no outside global/local variable shall work here) --> it fails.
  * your data must be available either in the code or any other viable way (like a public API)
  * package version and other session information can be shared
  * bonus : it handles sharing graphics with the *imgur* service
  
That way one could answer the question at hand with a complete set of information including the environment used (OS and packages version) as follow :

```{r complete and sharable answer, eval=FALSE, warning=FALSE, message=FALSE}
reprex::reprex(x = {
	dfr <- tibble::tribble(
		~var1,          ~var2, ~var3,
		"gr¦pefruit",   "ugli fruit",    4L,
		"huckleberry",      "kumquat",    6L,
		"duri¦n", "blackcurrant",    1L,
		"d¦te",   "cantaloup¦",    1L,
		"fig", "canary m¦lon",    6L,
		"s¦l¦l berry",     "rambutan",    5L
	)
	dfr
	
	# Method 1 - base R
	dfr_clean <- dfr
	dfr_clean$var1 <- gsub(pattern = "¦", replacement = "a", x = dfr$var1)
	dfr_clean[["var2"]] <- gsub(pattern = "¦", replacement = "e", x = dfr$var2)
	
	# Method 2 - tidyverse
	suppressPackageStartupMessages( library(dplyr) )
	suppressPackageStartupMessages( library(stringr) )
	dfr %>% mutate(var1 = str_replace_all(string = var1, pattern = "¦", replacement = "a"),
		       var2 = str_replace_all(string = var2, pattern = "¦", replacement = "e")
	)
}, si = TRUE, venue = "so")
```

Which gives the following in the "Viewer" pane for preview and is also copied to the clipboard :

<!-- language-all: lang-r -->


    dfr <- tibble::tribble(
        ~var1,          ~var2, ~var3,
        "gr¦pefruit",   "ugli fruit",    4L,
        "huckleberry",      "kumquat",    6L,
        "duri¦n", "blackcurrant",    1L,
        "d¦te",   "cantaloup¦",    1L,
        "fig", "canary m¦lon",    6L,
        "s¦l¦l berry",     "rambutan",    5L
    )
    dfr
    #> # A tibble: 6 x 3
    #>   var1        var2          var3
    #>   <chr>       <chr>        <int>
    #> 1 gr¦pefruit  ugli fruit       4
    #> 2 huckleberry kumquat          6
    #> 3 duri¦n      blackcurrant     1
    #> 4 d¦te        cantaloup¦       1
    #> 5 fig         canary m¦lon     6
    #> 6 s¦l¦l berry rambutan         5

    # Method 1 - base R
    dfr_clean <- dfr
    dfr_clean$var1 <- gsub(pattern = "¦", replacement = "a", x = dfr$var1)
    dfr_clean[["var2"]] <- gsub(pattern = "¦", replacement = "e", x = dfr$var2)

    # Method 2 - tidyverse
    suppressPackageStartupMessages( library(dplyr) )
    suppressPackageStartupMessages( library(stringr) )
    dfr %>% mutate(var1 = str_replace_all(string = var1, pattern = "¦", replacement = "a"),
               var2 = str_replace_all(string = var2, pattern = "¦", replacement = "e")
    )
    #> # A tibble: 6 x 3
    #>   var1        var2          var3
    #>   <chr>       <chr>        <int>
    #> 1 grapefruit  ugli fruit       4
    #> 2 huckleberry kumquat          6
    #> 3 durian      blackcurrant     1
    #> 4 date        cantaloupe       1
    #> 5 fig         canary melon     6
    #> 6 salal berry rambutan         5

<sup>Created on 2020-11-08 by the [reprex package](https://reprex.tidyverse.org) (v0.3.0)</sup>

    devtools::session_info()
    #> - Session info ---------------------------------------------------------------
    #>  setting  value                       
    #>  version  R version 4.0.2 (2020-06-22)
    #>  os       Windows 10 x64              
    #>  system   x86_64, mingw32             
    #>  ui       RTerm                       
    #>  language (EN)                        
    #>  collate  French_France.1252          
    #>  ctype    French_France.1252          
    #>  tz       Europe/Paris                
    #>  date     2020-11-08                  
    #> 
    #> - Packages -------------------------------------------------------------------
    #>  package     * version date       lib source        
    #>  assertthat    0.2.1   2019-03-21 [1] CRAN (R 4.0.2)
    #>  backports     1.1.7   2020-05-13 [1] CRAN (R 4.0.0)
    #>  callr         3.4.3   2020-03-28 [1] CRAN (R 4.0.2)
    #>  cli           2.0.2   2020-02-28 [1] CRAN (R 4.0.2)
    #>  crayon        1.3.4   2017-09-16 [1] CRAN (R 4.0.2)
    #>  desc          1.2.0   2018-05-01 [1] CRAN (R 4.0.2)
    #>  devtools      2.3.0   2020-04-10 [1] CRAN (R 4.0.2)
    #>  digest        0.6.25  2020-02-23 [1] CRAN (R 4.0.2)
    #>  dplyr       * 1.0.0   2020-05-29 [1] CRAN (R 4.0.2)
    #>  ellipsis      0.3.1   2020-05-15 [1] CRAN (R 4.0.2)
    #>  evaluate      0.14    2019-05-28 [1] CRAN (R 4.0.2)
    #>  fansi         0.4.1   2020-01-08 [1] CRAN (R 4.0.2)
    #>  fs            1.4.2   2020-06-30 [1] CRAN (R 4.0.2)
    #>  generics      0.0.2   2018-11-29 [1] CRAN (R 4.0.2)
    #>  glue          1.4.1   2020-05-13 [1] CRAN (R 4.0.2)
    #>  highr         0.8     2019-03-20 [1] CRAN (R 4.0.2)
    #>  htmltools     0.5.0   2020-06-16 [1] CRAN (R 4.0.2)
    #>  knitr         1.29    2020-06-23 [1] CRAN (R 4.0.2)
    #>  lifecycle     0.2.0   2020-03-06 [1] CRAN (R 4.0.2)
    #>  magrittr      1.5     2014-11-22 [1] CRAN (R 4.0.2)
    #>  memoise       1.1.0   2017-04-21 [1] CRAN (R 4.0.2)
    #>  pillar        1.4.6   2020-07-10 [1] CRAN (R 4.0.2)
    #>  pkgbuild      1.1.0   2020-07-13 [1] CRAN (R 4.0.2)
    #>  pkgconfig     2.0.3   2019-09-22 [1] CRAN (R 4.0.2)
    #>  pkgload       1.1.0   2020-05-29 [1] CRAN (R 4.0.2)
    #>  prettyunits   1.1.1   2020-01-24 [1] CRAN (R 4.0.2)
    #>  processx      3.4.3   2020-07-05 [1] CRAN (R 4.0.2)
    #>  ps            1.3.3   2020-05-08 [1] CRAN (R 4.0.2)
    #>  purrr         0.3.4   2020-04-17 [1] CRAN (R 4.0.2)
    #>  R6            2.4.1   2019-11-12 [1] CRAN (R 4.0.2)
    #>  remotes       2.1.1   2020-02-15 [1] CRAN (R 4.0.2)
    #>  rlang         0.4.7   2020-07-09 [1] CRAN (R 4.0.2)
    #>  rmarkdown     2.3     2020-06-18 [1] CRAN (R 4.0.2)
    #>  rprojroot     1.3-2   2018-01-03 [1] CRAN (R 4.0.2)
    #>  sessioninfo   1.1.1   2018-11-05 [1] CRAN (R 4.0.2)
    #>  stringi       1.4.6   2020-02-17 [1] CRAN (R 4.0.0)
    #>  stringr     * 1.4.0   2019-02-10 [1] CRAN (R 4.0.2)
    #>  testthat      2.3.2   2020-03-02 [1] CRAN (R 4.0.2)
    #>  tibble        3.0.3   2020-07-10 [1] CRAN (R 4.0.2)
    #>  tidyselect    1.1.0   2020-05-11 [1] CRAN (R 4.0.2)
    #>  usethis       1.6.1   2020-04-29 [1] CRAN (R 4.0.2)
    #>  utf8          1.1.4   2018-05-24 [1] CRAN (R 4.0.2)
    #>  vctrs         0.3.2   2020-07-15 [1] CRAN (R 4.0.2)
    #>  withr         2.2.0   2020-04-20 [1] CRAN (R 4.0.2)
    #>  xfun          0.15    2020-06-21 [1] CRAN (R 4.0.2)
    #>  yaml          2.2.1   2020-02-01 [1] CRAN (R 4.0.0)
    #> 
    #> [1] C:/Users/cbonamy/Documents/R/win-library/4.0
    #> [2] C:/Program Files/R/R-4.0.2/library



R References :

- [SO how-to-make-a-great-r-reproducible-example](https://stackoverflow.com/questions/5963269/how-to-make-a-great-r-reproducible-example)
- [R Community faq](https://community.rstudio.com/t/faq-whats-a-reproducible-example-reprex-and-how-do-i-do-one/5219)
- [R Community faq 2](https://community.rstudio.com/t/faq-how-to-do-a-minimal-reproducible-example-reprex-for-beginners/23061)


## To go further

One could share Docker container to fix the OS, packages version and so on. There are 2 nice package manager tool for that in Python and R with respectively `conda` (package management system) and `renv` (a R package that detect and register your configuration and used packages version).